Optimal management of a stochastically varying population when policy adjustment is costly

### 2014-05-26 manuscripts/manuscript.Rmd
start to tune plotting

```{r libraries, output="hide", include=FALSE, cache=FALSE}
rm(list=ls())
library("pdgControl")
library("reshape2")
library("ggplot2")
library("data.table")
library("pander")
library("cboettigR")
theme_set(theme_bw())
```

@@

Ecosystems are dynamic and exhibit rich patterns of variability in
both time and space. In designing management policies for ecosystems,
managers need to decide how much of that variation to respond to. Managers
could try to track variations in ecosystem dynamics very closely by
setting policies that are extremely responsive to the environment. Many
theoretical studies that seek to identify optimal policies for exploited
populations and communities adopt this approach and largely ignore the
challenges that would be involved in implementing such recommendations
[e.g., @Reed1979; @Neubert2003; @Sethi2005; @Halpern2011]. However, 
the policy process can often be much more sluggish
to respond to variations in ecosystem dynamics [@Walters1978; @Armsworth2012]. 
Moreover, stakeholders impacted by ecosystem management
may prefer some stability and not want to deal with continually changing
management recommendations [@Biais1995; @Armsworth2003;
@Patterson2007a; @Patterson2007b; @Sanchirico2008].
In other words, whatever gains are available from fine-tuning a
policy prescription to reflect environmental variation more closely
should be traded off against potential costs associated with the more
interventionist approach to management this would require.

@@

```{r Figure_1, fig.cap="Expected net present value by functional form of penalty. Horizontal axis shows the coefficient $c_i$ governing the magnitude of the policy cost, while vertical axis shows fraction of the maximum expected net present value dissipated by the policy cost.  The horizontal line indicates a value of the stock that is 80% of the maximum expected value in the absence of policy adjustment costs, $NPV_0({\\bf h_0^*})$. Selecting the coefficient $c_i$ corresponding to this value in each functional form allows us to make consistent comparisons across the different functional forms of policy costs.", dependson="quadcosts", fig.width=5, fig.height=3}
ggplot(fees, aes(c2, (npv0-value)/npv0, col=variable)) + geom_point() + geom_line()

@@

```{r Figure_2, dependson="tidy", fig.cap="Individual realizations of optimal harvesting strategy under the different functional forms of adjustment costs.", fig.height=3, fig.width=8}
ggplot(subset(dt, replicate %in% names(reps)[1:5])) +
  geom_line(aes(time, alternate), col="grey20", lwd=1) +
  geom_line(aes(time, fishstock), col=rgb(0,0,1,.8)) + 
  facet_grid(replicate~penalty_fn) + 
  labs(x="time", y="stock size", title = "Stock Dynamics") +
  theme_bw()

@@

```{r Figure_3, dependson="tidy", fig.cap="Individual realizations of optimal stock size strategy under the different functional forms of adjustment costs.", fig.height=3, fig.width=8}
ggplot(subset(dt, replicate %in% names(reps)[1:5])) +
  geom_line(aes(time, harvest_alt), col="grey20", lwd=1)  +
  geom_line(aes(time, harvest), col=rgb(0,0,1,.8)) + 
  facet_grid(replicate~penalty_fn) + 
  labs(x="time", y="havest intensity (fish taken)", title = "Harvest Policy Dynamics") + 
  theme_bw()

@@

```{r Figure_4, fig.cap="(a) Variance of optimal harvest rates through time as a function of size of adjustment penalties, the latter being calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs. (b) Autocorrelation of optimal harvest rates through time as a function of size of adjustment penalties, the latter being calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs.", fig.height=3, fig.width=8}
Fig4a <- ggplot(harvest_trends, 
                aes(fraction, Ev, ymin=Ev-SDv, ymax=Ev+SDv, col=penalty)) + 
  geom_ribbon(aes(fill=penalty, col=NA), lwd=0, alpha=.05) + 
  geom_line() + 
  xlab("Fraction of NPV lost to costs") +
  scale_x_continuous(limits=c(0, 0.4))

Fig4b <- ggplot(harvest_trends, 
                aes(fraction, Ea, ymin=Ea-SDa, ymax=Ea+SDa, col=penalty)) + 
  geom_ribbon(aes(fill=penalty, col=NA), lwd=0, alpha=.05) + 
  geom_line() + xlab("Fraction of NPV lost to costs")+
  scale_x_continuous(limits=c(0, 0.4))

@@

```{r Figure_5, fig.cap = "(a) Variance of optimal stock sizes through time as a function of size of adjustment penalties, the latter being calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs. (b) Autocorrelation of optimal stock sizes through time as a function of size of adjustment penalties, the latter being calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs.", fig.height=3, fig.width=8}
Fig5a <- ggplot(fishstock_trends, 
                aes(fraction, Ev, ymin=Ev-SDv, ymax=Ev+SDv, col=penalty)) + 
  geom_ribbon(aes(fill=penalty, col=NA), lwd=0, alpha=.05) + 
  geom_line() + xlab("Fraction of NPV lost to costs")+
  scale_x_continuous(limits=c(0, 0.4))

Fig5b <- ggplot(fishstock_trends, 
                aes(fraction, Ea, ymin=Ea-SDa, ymax=Ea+SDa, col=penalty)) + 
  geom_ribbon(aes(fill=penalty, col=NA), lwd=0, alpha=.05) + 
  geom_line() + xlab("Fraction of NPV lost to costs") +
  scale_x_continuous(limits=c(0, 0.4))

>>>>

```{r libraries, output="hide", include=FALSE, cache=FALSE}
rm(list=ls())
library("pdgControl")
library("reshape2")
library("ggplot2")
library("data.table")
library("pander")
library("cboettigR")
library("ggthemes")
theme_set(theme_tufte())
```

@@

Ecosystems are dynamic and exhibit rich patterns of variability in
both time and space. In designing management policies for ecosystems,
managers need to decide how much of that variation to respond to. Managers
could try to track variations in ecosystem dynamics very closely by
setting policies that are extremely responsive to the environment. Many
theoretical studies that seek to identify optimal policies for exploited
populations and communities adopt this approach and largely ignore the
challenges that would be involved in implementing such recommendations
[e.g., @Reed1979; @Neubert2003; @Sethi2005; @Halpern2011]. However, 
the policy process can often be much more sluggish
to respond to variations in ecosystem dynamics [@Walters1978; @Armsworth2010]. 
Moreover, stakeholders impacted by ecosystem management
may prefer some stability and not want to deal with continually changing
management recommendations [@Biais1995; @Armsworth2003;
@Patterson2007a; @Patterson2007b; @Sanchirico2008].
In other words, whatever gains are available from fine-tuning a
policy prescription to reflect environmental variation more closely
should be traded off against potential costs associated with the more
interventionist approach to management this would require.

@@

```{r Figure_1, fig.cap="Expected net present value by functional form of penalty. Horizontal axis shows the coefficient $c_i$ governing the magnitude of the policy cost, while vertical axis shows fraction of the maximum expected net present value dissipated by the policy cost.  The horizontal line indicates a value of the stock that is 80% of the maximum expected value in the absence of policy adjustment costs, $NPV_0({\\bf h_0^*})$. Selecting the coefficient $c_i$ corresponding to this value in each functional form allows us to make consistent comparisons across the different functional forms of policy costs.", dependson="quadcosts", fig.width=5, fig.height=3}
ggplot(fees, aes(c2, (npv0-value)/npv0, lty=variable)) + geom_line() + theme_tufte()

@@

```{r Figure_2, dependson="tidy", fig.cap="Individual realizations of optimal harvesting strategy under the different functional forms of adjustment costs.", fig.height=3, fig.width=8}
ggplot(subset(dt, replicate %in% names(reps)[1:5])) +
  geom_line(aes(time, alternate), col="grey20", lwd=1) +
  geom_line(aes(time, fishstock), col=rgb(0,0,1,.8)) + 
  facet_grid(replicate~penalty_fn) + 
  labs(x="time", y="stock size", title = "Stock Dynamics")  + theme_tufte()

@@

```{r Figure_3, dependson="tidy", fig.cap="Individual realizations of optimal stock size strategy under the different functional forms of adjustment costs.", fig.height=3, fig.width=8}
ggplot(subset(dt, replicate %in% names(reps)[1:5])) +
  geom_line(aes(time, harvest_alt), col="grey20", lwd=1)  +
  geom_line(aes(time, harvest), col=rgb(0,0,1,.8)) + 
  facet_grid(replicate~penalty_fn) + 
  labs(x="time", y="havest intensity (fish taken)", title = "Harvest Policy Dynamics")  + theme_tufte()

@@

```{r Figure_4, fig.cap="(a) Variance of optimal harvest rates through time as a function of size of adjustment penalties, the latter being calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs. (b) Autocorrelation of optimal harvest rates through time as a function of size of adjustment penalties, the latter being calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs.", fig.height=3, fig.width=8}
Fig4a <- ggplot(harvest_trends, 
                aes(fraction, Ev, ymin=Ev-SDv, ymax=Ev+SDv, col=penalty)) + 
  geom_ribbon(aes(fill=penalty, col=NA), lwd=0, alpha=.05) + 
  geom_line() + 
  xlab("Fraction of NPV lost to costs") +
  scale_x_continuous(limits=c(0, 0.4)) + theme_tufte()

Fig4b <- ggplot(harvest_trends, 
                aes(fraction, Ea, ymin=Ea-SDa, ymax=Ea+SDa, col=penalty)) + 
  geom_ribbon(aes(fill=penalty, col=NA), lwd=0, alpha=.05) + 
  geom_line() + xlab("Fraction of NPV lost to costs")+
  scale_x_continuous(limits=c(0, 0.4)) + theme_tufte()

@@

```{r Figure_5, fig.cap = "(a) Variance of optimal stock sizes through time as a function of size of adjustment penalties, the latter being calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs. (b) Autocorrelation of optimal stock sizes through time as a function of size of adjustment penalties, the latter being calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs.", fig.height=3, fig.width=8}
Fig5a <- ggplot(fishstock_trends, 
                aes(fraction, Ev, ymin=Ev-SDv, ymax=Ev+SDv, col=penalty)) + 
  geom_ribbon(aes(fill=penalty, col=NA), lwd=0, alpha=.05) + 
  geom_line() + xlab("Fraction of NPV lost to costs")+
  scale_x_continuous(limits=c(0, 0.4))  + theme_tufte()

Fig5b <- ggplot(fishstock_trends, 
                aes(fraction, Ea, ymin=Ea-SDa, ymax=Ea+SDa, col=penalty)) + 
  geom_ribbon(aes(fill=penalty, col=NA), lwd=0, alpha=.05) + 
  geom_line() + xlab("Fraction of NPV lost to costs") +
  scale_x_continuous(limits=c(0, 0.4))  + theme_tufte()
